---
title: "ggplot 1"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggridges)
```


## create the weatger data

```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## create a ggplot
```{r}
ggplot(weather_df, aes(x=tmin, y=tmax)) +
  geom_point()
##aes is for aesthetics
##geom_point() adds a corresponding point to each thing
```

alternate way of making this plot
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax))+geom_point()
```
saving initial points. Most of the time i do not do this
```{r}
scatterplot=
  weather_df %>%
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point()

#doesn't print out results. just saves things

scatterplot

#this prints it
```
time to add some color ... 
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(aes(color=name), alpha=0.4)
#alpha level determines transparency. closer to 0, mostly transparent
```
difference between aes in the geom_point vs in the ggplot. "why do 'aes' positions matter?

color matters in geom_point but not in geom smooth

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(aes(color=name), alpha=0.4) +
  geom_smooth(se=FALSE)

#se=false gets rid of the error bars
```

if put it in ggplot, we see that it goes through all of the points (individual)
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(aes(color=name), alpha=0.4) +
  geom_smooth(se=FALSE)

#se=false gets rid of the error bars
```

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(aes(color=name), alpha=0.4) +
  geom_smooth(se=FALSE) +
  facet_grid(~name)

#se=false gets rid of the error bars
#facet panels each group
```

more fun plots! 

```{r}
#seasonal trend in each of these locations
weather_df %>%
  ggplot(aes(x=date, y=prcp, color=name)) +
  geom_point() +
  geom_smooth(se=FALSE)

#geom_point is the signifier for scatterplots
#can change the y
```
```{r}
#seasonal trend in each of these locations
weather_df %>%
  ggplot(aes(x=date, y=tmax, color=name)) +
  geom_point(aes(size=prcp), alpha=0.5) +
  geom_smooth(size=2, se=FALSE)

#geom_point is the signifier for scatterplots
#prcp is changed too 
```
just the smooth curve 
```{r}
weather_df %>%
  ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```
2d density
```{r}

weather_df %>%
  ggplot(aes(x=tmin, y=tmax)) + 
  geom_bin2d()
```
# More kinds of plots!

##univariate plots

###histogram

```{r}
  ggplot(weather_df,aes(x = tmax, color=name)) + 
  geom_histogram()
```
```{r}
weather_df %>%
  ggplot(aes(x=tmax, fill=name)) + 
  geom_histogram(position="dodge", binwidth=2)
```

```{r}
weather_df %>%
  ggplot(aes(x=tmax, fill=name)) + 
  geom_histogram(position="dodge", binwidth=2) +
  facet_grid(~name)
```
### density plot

```{r}
weather_df %>%
  ggplot(aes(x=tmax, fill=name)) + 
  geom_density(alpha=0.3)
```

### boxplots

```{r}
weather_df %>%
  ggplot(aes(x=name, y=tmax, fill=name)) +
  geom_boxplot()
```

### violin plots 
```{r}
weather_df %>%
  ggplot(aes(x=name, y=tmax, fill=name)) +
  geom_violin()
```
### ridge plots
```{r}
weather_df %>%
  ggplot(aes(x=tmax, y=name)) +
  geom_density_ridges()
```

##saving a plot

```{r}
ggp_ridge_tmp=
  weather_df %>%
  ggplot(aes(x=tmax, y=name)) +
  geom_density_ridges()

ggsave("ggplot_temp_ridge.pdf", ggp_ridge_tmp)
```
##embedding plots in R Markdown

```{r fig.width=6}
weather_df %>%
  ggplot(aes(x=tmin,y=tmax, color=name)) +
  geom_point(alpha=0.4)+
  geom_smooth(se=FALSE)

```


