---
title: "qqplot 2"
date: "10/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggridges)
```


## create the weatger data

```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```
cache saves output in directory and doesnt redownload the dataset

## making new plots

start with an old plot!
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha=0.5)
```

add labels:
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha=0.5) + 
  labs(
    title = "Temperature plot",
    x="Minimum Temp (C)",
    y= "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  )
```
x axis and tick marks
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha=0.5) + 
  labs(
    title = "Temperature plot",
    x="Minimum Temp (C)",
    y= "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15,-5,20),
    labels = c("-15C", "-5" , "20")
  ) +
  scale_y_continuous(
    trans = "sqrt"
  )
```
breaks: currently -10, 0, 10, 20. 
(the width on the bottom)

trans= transformation. 
  scale_y_continuous(
    trans = "sqrt" 
  ) and scale_y_sqrt() are the same function in r. can use either

## colors!
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha=0.5) + 
  labs(
    title = "Temperature plot",
    x="Minimum Temp (C)",
    y= "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_color_hue(
    name = "Weather Station Location",
    h=c(50,200)
  )
```

name = the legend name
you can change the color by hand but you do not have to do this

###viridis package
```{r}
ggplot_base=
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name))+
  geom_point(alpha=0.5) + 
  labs(
    title = "Temperature plot",
    x="Minimum Temp (C)",
    y= "Maximum Temp (C)",
    caption = "Data from NOAA via rnoaa package"
  ) +
 viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggplot_base

```
##themes
```{r}
ggplot_base +
  theme(legend.position="bottom")
```

theme: global things that describe what hte plot looks like 
- general organization (i.e.legend position) **better to put legends on the bottom**

```{r}
ggplot_base +
  theme_bw()+
  theme(legend.position="bottom")
```
how to get rid of the legend in total
```{r}
ggplot_base +
  theme_bw()+
  theme(legend.position="none")
```

order matters. theme minimal and tehn update positions, its going to go in order
```{r}
ggplot_base +
  theme_minimal()+
  theme(legend.position="bottom")
```
 viridis overrides some default color pallet. This is just so that all plots look the smae. 
```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## more than one dataset
```{r}
central_park=
  weather_df %>% 
  filter (name=="CentralPark_NY")

waikiki=
  weather_df %>% 
  filter(name=="Waikiki_HA")

  ggplot(data=waikiki, aes(x=date, y=tmax, color=name)) +
  geom_point(aes(size=prcp))+
    geom_line(data=central_park)
```

(brief aside about colors)

```{r}
waikiki %>% 
  ggplot(aes(x= date, y=tmax)) + 
  geom_point(color= "yellow")
```
devtools::install_github("thomasp85/patchwork")

##multi=panel plots
```{r}
library(patchwork)
ggp_scatter=
  weather_df %>% 
  ggplot(aes(x=tmin, y=tmax)) + 
  geom_point()

ggp_density=
  weather_df %>%  
  ggplot(aes(x=tmin)) +
  geom_density()

ggp_box=
  weather_df %>% 
  ggplot(aes(x = name, y =tmax)) + 
  geom_boxplot()

ggp_scatter + (ggp_density / ggp_box)
```
## Data Manipulation

```{r}
weather_df %>% 
  mutate(
    name=factor(name),
    name=fct_relevel(name, "Waikiki_HA", "CentralPark_NY")
  ) %>% 
  ggplot(aes(x=name, y=tmax, color=name)) + 
  geom_boxplot()
```
relevel : puts them in different levels 
fct_reorder : puts them in order based on a variable 
```{r}
weather_df %>% 
  mutate(
    name=factor(name),
    name=fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x=name, y=tmax, color=name)) + 
  geom_boxplot()
```

## restructure the plot

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to="observation", 
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x=temperature, fill = observations)) +
  geom_density() + 
  feom_desnity() +
  facet_grid(-name) +
  theme(legend.position = "bottom")

```

